
---
title: "All Australia Benzo and Opioid Basic Stats"
author:  "Mofi Islam and Dennis Wollersheim "
date: "2018-04-30"
output:
workflowr::wflow_html:
toc: false
---

```{r pre_initial, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

rm(list=ls())
options(width = 200)
show_code = FALSE
source("lib/functions.R")
source("lib/mapping_functions.R")
source("lib/mapping_paper.R")
source("lib/standardisation.R")
source("lib/get_data.R")
source("lib/generate_data_frames.R")
library('tidyverse')
opts_chunk$set(echo = FALSE)
opts_chunk$set(fig.width=12, fig.height=8) 

```


```{r initial, results='hide', message=FALSE, warning=FALSE}

dataset="_rr"
dataset=""


get_data_from_cache(dataset = dataset)

df  %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  group_by(  pin, sex, age, drug_type, lga, state   ) %>%
  summarise(n_dose = sum( n_dose ) )%>%
  spread( drug_type, n_dose ) %>%
  rename( n_dose_opioid = opioid, 
         n_dose_benzo = benzodiazepine ) %>%
  mutate_if( is_numeric, function(x) { ifelse( is.na( x ), 0, x ) } ) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( urbanization7 = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  ), 
                      ifelse( urbanization7=='Rural Agriculture', 'Rural Agricultural', urbanization7)) %>%
               rename( urbanization22 = class_name ) %>%
               rename( urbanization2 = urbanization ) %>%
               group_by( lga, lga_name, seifa, urbanization2, urbanization7, urbanization22  ) %>%
               summarise( population_2013 = sum( population )), 
             by=c('lga') ) %>% 
  inner_join( df_patient_usage, ) %>%
  ungroup() %>%
  { . } -> df_mlm

```
# Methods

The data is from the PBS prescribing data.  In this dataset, we have every opioid and benzodiazepine prescription for a 10% sample of the population, made in the period 2013-2016.  In a single prescription, we have a date of supply, a drug type and strength, and the number of pills.  Using the WHO DDD drug classification, we determine the total DDD for each prescription, and then we estimate the end day of the prescription to be the start day + totalDDD, rounded to the nearest day.  This start and end data comprise the prescription period. 

## Variables

### Independent
* LGA level variables
  * SEIFA
  * urbanisation
  * census data - english language, 2 person household
* Person level variables
  * gender
  * age groups
  * drug of choice
  * user type
  * lga
* Script level variables
  * month of supply
  * drug type - drug name
  * benzo type
  * opioid type


### dependent variables
* benzo and opioid usage (scripts / ddd / number of users / ddd per user / scripts per user  )
* lga level ddd
* usage category


# Number of users using for number of quarters, for both drug categories

```{r nquarter }

df_mlm %>%
  gather(type, nquarter, opioid_n_quarter, benzo_n_quarter) %>%
  filter( nquarter != 0 ) %>%
  count( type, nquarter ) %>%
  spread( type, n) %>% 
  mutate_at( vars( ends_with( 'n_quarter' )), funs(  pct=round( ./sum(.)*100,2))) %>%
  { . } -> df_nquarter

df_nquarter %>% kable()


df_nquarter %>%
  select( -ends_with('n_quarter')) %>%
  gather( type, pct, -nquarter )  %>%
  ggplot( aes( pct,nquarter, color=type )) + 
  geom_point() + 
  geom_path()


```
# Percentages of of users within drug category by age group

```{r quarters_by_age}

df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( opioid_usage_category, age ) %>%
  filter( !is.na( opioid_usage_category )) %>%
  group_by(opioid_usage_category ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( color=opioid_usage_category, pct, x=age, group=opioid_usage_category )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'opioid number of users percentage of each usage category by age group ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )



df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( benzo_usage_category, age ) %>%
  filter( !is.na( benzo_usage_category )) %>%
  group_by(benzo_usage_category ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( color=benzo_usage_category, pct, x=age, group=benzo_usage_category )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'benzo number of users percentage of each  usage category by age group ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )


df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( both_category, age ) %>%
  filter( !is.na( both_category )) %>%
  group_by(both_category ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( color=both_category, pct, x=age, group=both_category )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'both Opioid and benzo number of users percentage of each  usage category by age group ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )


```


# Number of users percentages of age group by usage category

```{r quarters_by_age_v1}

df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( opioid_usage_category, age ) %>%
  filter( !is.na( opioid_usage_category )) %>%
  group_by(age ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( opioid_usage_category, pct, color=age, group=age )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'opioid percentage of each age group by  usage category ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )


df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( benzo_usage_category, age ) %>%
  filter( !is.na( benzo_usage_category )) %>%
  group_by(age ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( benzo_usage_category, pct, color=age, group=age )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'benzo percentage of each age group by  usage category ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )


df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( both_category, age ) %>%
  filter( !is.na( both_category )) %>%
  group_by(age ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

df_tograph %>%
  ggplot(  aes( both_category, pct, color=age, group=age )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'both opioid  and benzo percentage of each age group by  usage category ')

df_tograph %>%
  spread( age, pct ) %>% 
  kable( )

```

# Number of users  percentages of usage category by state for each age group

** this should be standardized because population distribution varies so much between states

```{r category_by_state}

df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( opioid_usage_category, age, state ) %>%
  filter( !is.na( opioid_usage_category )) %>%
  group_by(opioid_usage_category, state ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  arrange(opioid_usage_category, state ) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

#df_tograph %>%
#  spread( age, pct ) %>% 
#  kable( )

df_tograph %>%
  ggplot(  aes( opioid_usage_category, pct, color=age, group=age )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'opioid usate category vs % of this category made up by this age group by age group and state') +
  facet_wrap( ~state )


df_mlm %>%
  select( age, state, ends_with('category')) %>%
  count( benzo_usage_category, age, state ) %>%
  filter( !is.na( benzo_usage_category )) %>%
  group_by(benzo_usage_category, state ) %>%
  mutate( pct = round( n / sum( n ) * 100, 2 )) %>%
  arrange(benzo_usage_category, state ) %>%
  select(-n ) %>%
  ungroup() %>%
  { . } -> df_tograph

#df_tograph %>%
#  spread( age, pct ) %>% 
#  kable( )

df_tograph %>%
  ggplot(  aes( benzo_usage_category, pct, color=age, group=age )) + 
  geom_point() + 
  geom_path() +
  ggtitle( ' usate category vs % of this category made up by this age group by age group and state') +
  facet_wrap( ~state )

```

# ddd of each drug type by state  and usage category

```{r ddd_and_type_by_state_usage}

df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, opioid_usage_category, state, type_name, supply_year )) %>% 
  group_by( opioid_usage_category, state, type_name ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph

#df_tograph %>%
#  spread( type_name, ddd ) %>% 
#  kable( )

df_tograph %>%
  ungroup() %>%
  ggplot(  aes( opioid_usage_category, ddd, color=type_name, group=type_name )) + 
  geom_point() + 
  geom_path() +
  ggtitle( 'ddd for each usage category by drug_type and state') +
  facet_wrap( ~state )



```

# ddd per month for each state and age group
state mean from lga level standardisation

```{r ddd_and_category_by_state}

df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, opioid_usage_category, state, supply_month )) %>% 
  group_by( opioid_usage_category, state, supply_month ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph

#df_tograph %>%
#  spread(opioid_usage_category, ddd ) %>% 
#  kable( )

df_tograph %>%
  ungroup() %>%
  mutate( dom = as.Date( paste0(supply_month, '-01'), format='%Y-%m-%d')) %>%
  ggplot(  aes( dom , ddd, color=state, group=state)) + 
  geom_path() +
  ggtitle( 'ddd for each opioid usage category by drug_type and state') +
  scale_x_date(date_breaks = "6 months" , date_labels = "%Y-%m") +
  facet_wrap( ~opioid_usage_category )



df %>%
  filter( drug_type=='benzodiazepine' ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, benzo_usage_category, state, supply_month )) %>% 
  group_by( benzo_usage_category, state, supply_month ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph

#df_tograph %>%
#  spread(benzo_usage_category, ddd ) %>% 
#  kable( )

df_tograph %>%
  ungroup() %>%
  mutate( dom = as.Date( paste0(supply_month, '-01'), format='%Y-%m-%d')) %>%
  ggplot(  aes( dom , ddd, color=state, group=state)) + 
  geom_path() +
  ggtitle( 'ddd for each benzo usage category by drug_type and state') +
  scale_x_date(date_breaks = "6 months" , date_labels = "%Y-%m") +
  facet_wrap( ~benzo_usage_category )



```
# ddd per month for each state and drug_type
state mean from lga level standardisation

```{r ddd_and_type_by_state}

df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, type_name, state, supply_month )) %>% 
  group_by( type_name, state, supply_month ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph

#df_tograph %>%
#  spread(type_name, ddd ) %>% 
#  kable( )

df_tograph %>%
  ungroup() %>%
  mutate( dom = as.Date( paste0(supply_month, '-01'), format='%Y-%m-%d')) %>%
  ggplot(  aes( dom , ddd, color=state, group=state)) + 
  geom_path() +
  ggtitle( 'ddd for each usage category by drug_type and state') +
  scale_x_date(date_breaks = "6 months" , date_labels = "%Y-%m") +
  facet_wrap( ~type_name, scales = 'free'  )


df %>%
  filter( drug_type=='benzodiazepine' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, type_name, state, supply_month )) %>% 
  group_by( type_name, state, supply_month ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph

#df_tograph %>%
#  spread(type_name, ddd ) %>% 
#  kable( )

df_tograph %>%
  ungroup() %>%
  mutate( dom = as.Date( paste0(supply_month, '-01'), format='%Y-%m-%d')) %>%
  ggplot(  aes( dom , ddd, color=state, group=state)) + 
  geom_path() +
  ggtitle( 'ddd for each usage category by drug_type and state') +
  scale_x_date(date_breaks = "6 months" , date_labels = "%Y-%m") +
  facet_wrap( ~type_name, scales = 'free'  )


```





