
---
title: "All Australia Benzo and Opioid Basic Stats"
author:  "Mofi Islam and Dennis Wollersheim "
date: "2018-04-30"
output:
workflowr::wflow_html:
toc: false
---

```{r pre_initial, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

rm(list=ls())
options(width = 200)
show_code = FALSE
source("lib/functions.R")
source("lib/mapping_functions.R")
source("lib/mapping_paper.R")
source("lib/standardisation.R")
source("lib/get_data.R")
source("lib/generate_data_frames.R")
library('tidyverse')
opts_chunk$set(echo = FALSE)
opts_chunk$set(fig.width=12, fig.height=8) 

```


```{r initial, results='hide', message=FALSE, warning=FALSE}

dataset="_rr"
dataset=""


get_data_from_cache(dataset = dataset)

df  %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  group_by(  pin, sex, age, drug_type, lga, state   ) %>%
  summarise(n_dose = sum( n_dose ) )%>%
  spread( drug_type, n_dose ) %>%
  rename( n_dose_opioid = opioid, 
         n_dose_benzo = benzodiazepine ) %>%
  mutate_if( is_numeric, function(x) { ifelse( is.na( x ), 0, x ) } ) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( urbanization7 = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  ), 
                      ifelse( urbanization7=='Rural Agriculture', 'Rural Agricultural', urbanization7)) %>%
               rename( urbanization22 = class_name ) %>%
               rename( urbanization2 = urbanization ) %>%
               group_by( lga, lga_name, seifa, urbanization2, urbanization7, urbanization22  ) %>%
               summarise( population_2013 = sum( population )), 
             by=c('lga') ) %>% 
  inner_join( df_patient_usage, ) %>%
  ungroup() %>%
  { . } -> df_mlm

```

# state level stats for each opioid usage category and type name
  -  number of users, ddd, dddU  

```{r opioid_type_name_usage_cateogry}

df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd_users( standardise_over = qc(lga, opioid_usage_category, state, type_name, supply_year )) %>% 
  group_by( opioid_usage_category, state, type_name ) %>% 
  summarise( ddd_user = mean( ddd_user ) ) %>% 
  { . } -> df_tograph



df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, opioid_usage_category, state, type_name, supply_year )) %>% 
  group_by( opioid_usage_category, state, type_name ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph_ddd

df_tograph %>%
  spread( type_name, ddd_user ) %>% 
  kable()

ggplot() + 
  geom_line( data=df_tograph, mapping=aes( type_name, ddd_user,  color=state, group=state ))+
  facet_wrap( ~ opioid_usage_category)


ggplot() + 
  geom_line( data=df_tograph, mapping=aes( opioid_usage_category, ddd_user,  color=state, group=state ))+
  facet_wrap( ~ type_name)

cat('## comparision with regular ddd')
ggplot() + 
  geom_line( data=df_tograph_ddd, mapping=aes( type_name, ddd,  color=state, group=state ))+
  facet_wrap( ~ opioid_usage_category)


ggplot() + 
  geom_line( data=df_tograph_ddd, mapping=aes( opioid_usage_category, ddd,  color=state, group=state ))+
  facet_wrap( ~ type_name)


```

## who is responsible for the hydromorphone in QLD?




```{r qld_hydromorphone, results='asis'}


df %>%
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  filter( type_name=='Hydromorphone' ) %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( state == 'QLD' ) %>% 
  { . } -> a

cat('summary stats for qld hydromorphone;  number of doses, number of pp')
a %>% summarise( sum( n_dose ), n_distinct( pin )) %>% kable()

cat('top 10 users')
a %>% 
  group_by( pin ) %>% 
  summarise( n_dose = sum( n_dose ), n()) %>%
  arrange( desc( n_dose )) %>% 
  head(10) %>% 
  kable()

cat(' top user')
df %>% 
  filter( pin == '7538740161') %>% 
  filter( drug_type=='opioid' & type_name != 'Other'  ) %>%
  arrange( supply_date ) %>% 
  kable

```

# state level stats for each benzo usage category and type name
  -  number of users, ddd, dddU  

```{r benzo_type_name_usage_cateogry}

df %>%
  filter( drug_type=='benzodiazepine' & type_name != 'Other'  ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd_users( standardise_over = qc(lga, benzo_usage_category, state, type_name, supply_year )) %>% 
  group_by( benzo_usage_category, state, type_name ) %>% 
  summarise( ddd_user = mean( ddd_user ) ) %>% 
  { . } -> df_tograph


df %>%
  filter( drug_type=='benzodiazepine' ) %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_patient_usage, by='pin'  ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  select_and_standardise_ddd( standardise_over = qc(lga, benzo_usage_category, state, type_name, supply_year )) %>% 
  group_by( benzo_usage_category, state, type_name ) %>% 
  summarise( ddd = mean( ddd ) ) %>% 
  { . } -> df_tograph_ddd

df_tograph %>%
  spread( type_name, ddd_user ) %>% 
  kable()

ggplot() + 
  geom_line( data=df_tograph, mapping=aes( type_name, ddd_user,  color=state, group=state ))+
  facet_wrap( ~ benzo_usage_category)

ggplot() + 
  geom_line( data=df_tograph, mapping=aes( benzo_usage_category, ddd_user,  color=state, group=state ))+
  facet_wrap( ~ type_name, scales='free_y')

cat('## comparision with regular ddd')

ggplot() + 
  geom_line( data=df_tograph_ddd, mapping=aes( type_name, ddd,  color=state, group=state ))+
  facet_wrap( ~ benzo_usage_category)

ggplot() + 
  geom_line( data=df_tograph_ddd, mapping=aes( benzo_usage_category, ddd,  color=state, group=state ))+
  facet_wrap( ~ type_name, scales='free_y')


```

# ddd_user exploration - comparison with various metrics, dose, users, and scripts per population 

  - % of users vs ddd_user
  - # scripts / population vs ddd_user
```{r explore_ddd_user}

df %>%
    inner_join( df_patient, by='pin' ) %>%
    filter( lga != '.' & !endsWith( lga,'99') ) %>%
    select_and_standardise_ddd_users( standardise_over = qc(lga, age, state, supply_year )) %>% 
    { . } -> df_tograph_ddd_user


df %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( lga != '.' & !endsWith( lga,'99') ) %>%
  group_by( lga, sex, age, state, supply_year ) %>% 
  summarise( n_dose = sum( n_dose ), n_scripts=n(), n_users=n_distinct( pin )) %>%
  ungroup() %>%
  f_join_population( rollup_level=qc(lga, state, supply_year, age) ) %>%
  mutate( 
          pct_dose = n_dose / population, 
          pct_scripts = n_scripts / population, 
          pct_users = n_users / population ) %>% 
  { . } -> df_stats


df_stats %>% 
  inner_join( df_tograph_ddd_user ) %>%
  ggplot( aes( pct_users, ddd_user, color=age ) ) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10()

df_stats %>% 
  inner_join( df_tograph_ddd_user ) %>%
  ggplot( aes( pct_dose, ddd_user, color=age ) ) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10()

df_stats %>% 
  inner_join( df_tograph_ddd_user ) %>%
  ggplot( aes( pct_scripts, ddd_user, color=age ) ) +
  geom_point() + 
  scale_x_log10() +
  scale_y_log10()



```

