---
title: "Benzo-opioid v2 analysis"
author:  "Mofi Islam and Dennis Wollersheim "
date: "2018-04-30"
output:
workflowr::wflow_html:
toc: false
---

```{r pre_initial, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

rm(list=ls())
options(width = 200)
show_code = FALSE
source("lib/functions.R")
source("lib/mapping_functions.R")
source("lib/standardisation.R")
source("lib/get_data.R")
source("lib/generate_data_frames.R")
library('tidyverse')
opts_chunk$set(echo = FALSE)

```

```{r initial, results='hide', message=FALSE, warning=FALSE}

dataset="_rr"
dataset=""


get_data_from_cache(dataset = dataset)


```


```{r multilevel_modelling}


df  %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  group_by(  pin, sex, age, drug_type, lga, state   ) %>%
  summarise(n_dose = sum( n_dose ) )%>%
  spread( drug_type, n_dose ) %>%
  rename( n_dose_opioid = opioid, 
         n_dose_benzo = benzodiazepine ) %>%
  mutate_if( is_numeric, function(x) { ifelse( is.na( x ), 0, x ) } ) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( urbanization7 = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  ), 
                      ifelse( urbanization7=='Rural Agriculture', 'Rural Agricultural', urbanization7)) %>%
               rename( urbanization22 = class_name ) %>%
               rename( urbanization2 = urbanization ) %>%
               group_by( lga, lga_name, seifa, urbanization2, urbanization7, urbanization22  ) %>%
               summarise( population_2013 = sum( population )), 
             by=c('lga') ) %>% 
 inner_join( df_patient_usage, ) %>%
 ungroup() %>%
  { . } -> df_mlm

foreign::write.dta(df_mlm, '/store/aarnet_owncloud/Sharing_between_mofi_and_dennis/Opioid/Opioid/new_vs_continuing/AllAustralia.dta')


df_mlm %>% 
  filter( drug_type=='opioid' ) %>%
  ungroup() %>%
  mutate( 
         sex = factor( sex ), 
         age = factor( age ), 
         lga = factor( lga ), 
         state= factor( state), 
         class_name = factor( class_name ), 
         class_subset = factor( class_subset ), 
         seifa = factor( seifa ), 
         urbanization = factor( urbanization )
         ) %>% 
  { . } -> df_mlm_opioid

df_mlm %>% 
  filter( drug_type!='opioid' ) %>%
  ungroup() %>%
	mutate( 
       sex = factor( sex ), 
       age = factor( age ), 
       lga = factor( lga ), 
       state= factor( state), 
	   class_name = factor( class_name ), 
	   class_subset = factor( class_subset ), 
       seifa = factor( seifa ), 
       urbanization = factor( urbanization )
 ) %>% 
{ . } -> df_mlm_benzo

df_mlm_opioid %>%
  lme4::lmer( n_dose ~ age + sex + seifa + (1|urbanization)  + (1|class_subset) + (1|class_name) + (1|lga), 
             data=. )->d

df_mlm_opioid %>%
  lme4::lmer( n_dose ~ age + sex + seifa + urbanization + class_subset + (1|state) + (1|lga), 
             data=. )->d


df_mlm_benzo %>%
   lme4::lmer( n_dose ~ age + sex + seifa + urbanization + class_subset + (1|state) + (1|lga), 
              data=. )->d


exp( lme4::fixef( d ))

d
