---
title: "Benzo-opioid v2 analysis"
author:  "Mofi Islam and Dennis Wollersheim "
date: "2018-04-30"
output:
workflowr::wflow_html:
toc: false
---

```{r pre_initial, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

rm(list=ls())
options(width = 200)
show_code = FALSE
source("lib/functions.R")
source("lib/mapping_functions.R")
source("lib/standardisation.R")
source("lib/get_data.R")
source("lib/generate_data_frames.R")
library('tidyverse')
opts_chunk$set(echo = FALSE)

```


```{r initial, results='hide', message=FALSE, warning=FALSE}

dataset="_rr"
dataset=""


get_data_from_cache(dataset = dataset)


```


Indeed, Queensland has higher ddd than the rest of the country (barring tasmania)
```{r explore , warning=FALSE}

df %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_population %>% distinct( lga, seifa, urbanization), by='lga' ) %>%
  filter( lga != '.' ) %>%
  select_and_standardise_ddd( standardise_over = qc( state ))

```
why are there 2 bumps in queensland?

```{r queensland, warning=FALSE  }

df %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_population %>% distinct( lga, seifa, urbanization, lga_name), by='lga' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  select_and_standardise_ddd( standardise_over = qc( state, lga )) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( class_subset = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  )) %>%
               filter( class_type != 'UCC' & class_type != 'RSG') %>%
               group_by( lga, lga_name, seifa, class_name, class_subset, urbanization  ) %>%
               summarise( population = sum( population )), 
             by=c('lga') ) %>% 
  { . } -> df_lga_ddd

df %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_population %>% distinct( lga, seifa, urbanization, lga_name), by='lga' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  select_and_standardise_ddd( standardise_over = qc( state, lga , drug_type )) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( class_subset = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  )) %>%
               filter( class_type != 'UCC' & class_type != 'RSG') %>%
               group_by( lga, lga_name, seifa, class_name, class_subset, urbanization  ) %>%
               summarise( population = sum( population )), 
             by=c('lga') ) %>% 
  { . } -> df_lga_ddd_drug_type
 

my_db_get_query("select  lga, supply_year, value as lote from  abs_census where measure_id = 'CENSUS_33'", 'mofi') %>%
  as.tibble() %>% 
  { . } -> df_lote

my_db_get_query("select  lga, supply_year, value as gini from  abs_census where measure_id = 'INCOME_41'", 'mofi') %>%
    as.tibble() %>% 
    { . } -> df_gini

qplot( lote, data=df_lote, log='x') -sadgg

df_lga_ddd %>%
  inner_join( filter( df_lote, supply_year==2011 ), by='lga') %>%
  arrange( desc( lote ))


df_lga_ddd %>%
  mutate( b=scale( ddd )) %>%
  ggplot( aes( x=b, fill=factor( seifa)) ) + 
  geom_dotplot(stackgroups=TRUE )


```

range of lga total population adjusted ddd

what do we want to know?



```{r test}


df_lga_ddd %>%
  inner_join( filter( df_lote, supply_year==2011 ), by='lga') %>%
  ggplot( aes( ddd, lote, color=urbanization )) +
  geom_smooth(method = "lm") +
  scale_x_log10() +
  scale_y_log10() +
  geom_point() - sadgg

df_lga_ddd %>% inner_join( filter( df_gini, supply_year==2011 ), by='lga') %>%
    ggplot( aes( ddd, gini, color=seifa )) +
    geom_smooth(method = "lm") +
    scale_x_log10() +
    scale_y_log10() +
    geom_point()  -sadgg


```


What is happening in QLD
Why is QLD different?

what aspects are QLD different

QLD urban is different
 - many remote areas
  - urban regional very large  - consistently high 

urbanization x state

```{r state_urbanization, echo=FALSE, fig.cap="state x urbanization", out.width = '100%'}
 


df_lga_ddd %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_name)) -sadgg

df_lga_ddd %>%
  filter( startsWith( as.character( urbanization), 'R')) %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_name)) + 
  facet_wrap(~class_name)  -sadgg

df_lga_ddd %>%
  filter( startsWith( as.character( urbanization), 'U')) %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_name)) + 
  facet_wrap(~class_name)  -sadgg


df_lga_ddd_drug_type %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_name), size=.1) + 
  facet_wrap( ~drug_type ) -sadgg

df_lga_ddd_drug_type %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_name), size=.1) + 
  facet_grid( urbanization~drug_type ) -sadgg


df_lga_ddd_drug_type %>%
  ggplot(aes(urbanization, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_subset), size=.1) + 
  facet_grid( drug_type ~ state ) +
  ggtitle( 'benzo and opioid comparison, country vs city' ) -sadgg 

df_lga_ddd_drug_type %>%
  filter( startsWith( as.character( urbanization), 'R')) %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_subset), size=.1) + 
  facet_grid(drug_type ~class_subset)  -sadgg

df_lga_ddd_drug_type %>%
  filter( startsWith( as.character( urbanization), 'U')) %>%
  ggplot(aes(state, ddd)) +
  geom_violin(colour = "grey50", mapping=aes(weight=population)) +
  geom_jitter( aes( color = class_subset)) + 
  facet_wrap(~class_subset)  -sadgg


df_population %>%
distinct( lga_name, class_name, state ) %>%
filter( state=='VIC'  ) %>%
arrange( class_name ) %>% View





df %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_population %>% 
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( class_subset = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  )) %>%
               distinct( lga, lga_name, seifa, class_name, class_subset, urbanization  ) ,
             by=c('lga') ) %>% 
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  select_and_standardise_ddd( standardise_over = qc( state, class_subset )) %>% 
  { . } -> df_state_class_ddd


df_state_class_ddd %>%
  ggplot(aes(state, ddd, fill=class_subset)) +
  geom_col() 

df %>%
  inner_join( df_patient, by='pin' ) %>%
  inner_join( df_population %>% 
             filter( lga != '.' & !endsWith( lga, '99') ) %>%
             mutate( class_subset = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  )) %>%
             distinct( lga, lga_name, seifa, class_name, class_subset, urbanization  ) ,
           by=c('lga') ) %>% 
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  select_and_standardise_ddd( standardise_over = qc( state, class_subset )) %>% 
  { . } -> df_state_class_ddd

df %>%
  inner_join( df_patient, by='pin' ) %>%
  filter( lga != '.' & !endsWith( lga, '99') ) %>%
  select_and_standardise_ddd( standardise_over = qc( state, lga )) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & !endsWith( lga, '99') ) %>%
               mutate( class_subset = str_replace( class_name,'([^ ]* [^ ]*).*',"\\1"  )) %>%
               filter( class_type != 'UCC' & class_type != 'RSG') %>%
               group_by( lga, lga_name, seifa, class_name, class_subset, urbanization,area_albers_sqkm ) %>%
               summarise( population = sum( population )) %>%
               mutate ( pp_sq_km = population / area_albers_sqkm), 
             by=c('lga') ) %>% 
  { . } -> df_lga_ddd_sqkm

df_lga_ddd_sqkm %>%
  ggplot(aes(pp_sq_km, ddd, color=class_subset)) +
  scale_x_log10() +
  geom_point()  + 
  geom_smooth( aes( pp_sq_km, ddd, color='1'))
```

Look at individual totals per year

