---
title: "Benzo-opioid v2 analysis"
author:  "Mofi Islam and Dennis Wollersheim "
date: "2018-04-30"
output:
workflowr::wflow_html:
toc: false
---

```{r pre_initial, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

rm(list=ls())
options(width = 200)
show_code = FALSE
source("lib/functions.R")
source("lib/mapping_functions.R")
source("lib/standardisation.R")
source("lib/get_data.R")
source("lib/generate_data_frames.R")
library('tidyverse')
opts_chunk$set(echo = FALSE)

```


```{r initial, results='hide', message=FALSE, warning=FALSE}

dataset="_rr"
dataset=""


get_data_from_cache(dataset = dataset)


```
# Methods

The data is from the PBS prescribing data.  In this dataset, we have every opioid and benzodiazepine prescription for a 10% sample of the population, made in the period 2013-2016.  In a single prescription, we have a date of supply, a drug type and strength, and the number of pills.  Using the WHO DDD drug classification, we determine the total DDD for each prescription, and then we estimate the end day of the prescription to be the start day + totalDDD, rounded to the nearest day.  This start and end data comprise the prescription period. 

## Variables

### Independent
* LGA level variables
  * SEIFA
  * urbanisation
  * census data - english language, 2 person household
* Person level variables
  * gender
  * age groups
  * drug of choice
  * user type
  * lga
* Script level variables
  * month of supply
  * drug type - drug name
  * benzo type
  * opioid type


### dependent variables
* benzo and opioid usage (scripts / ddd / number of users / ddd per user / scripts per user  )
* lga level ddd


```{r individual_level_graphs, message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient ) %>%
  group_by( supply_month, sex,  ) %>%
  summarise( n_dose = sum( n_dose ) ) %>%
  ggplot( aes( supply_month, n_dose, fill=sex )) +
  geom_col( position='dodge')

df %>%
  inner_join( df_patient ) %>%
  group_by( supply_month, sex,  ) %>%
  summarise( n = n()) %>%
  ggplot( aes( supply_month, n, fill=sex )) +
  geom_col( position='dodge')


df %>%
  inner_join( df_patient ) %>%
  group_by( supply_month, sex ) %>%
  summarise( n = n()) %>%
  ggplot( aes( supply_month, n, group=sex, color=sex )) +
  geom_line( ) 

```


```{r figure_103, message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient, by='pin') %>%
  filter( lga != '.' ) %>%
  select_and_standardise_ddd( standardise_over = qc( lga )) %>% 
  { . } -> base_figure_103

base_figure_103 %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               group_by( lga, state ) %>%
               summarise( population = sum( population )), 
              by='lga' ) %>% 
filter( !endsWith(lga,  '99' )) %>% 
{ . } -> figure_103


figure_103 %>%
  mutate( ddd_level = cut( ddd, breaks=0:30*5 )) %>%
  ggplot( aes( x=ddd_level, y=population, fill=state  )) +
  geom_col() +
  ggtitle( 'Total Australian population at various DDD levels')


```


```{r figure_103_age, message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient, by='pin') %>%
  filter( lga != '.' & age!='0-19' ) %>%
  select_and_standardise_ddd( standardise_over = qc( lga, age )) %>% 
  { . } -> base_figure_103_age

base_figure_103_age %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' & age!='0-19' ) %>%
               group_by( lga, state, age ) %>%
               summarise( population = sum( population )), 
              by=c('lga', 'age') ) %>% 
filter( !endsWith(lga,  '99' )) %>% 
{ . } -> figure_103_age


figure_103_age %>%
  mutate( ddd_level = cut( ddd, breaks=0:60*2 )) %>%
  ggplot( aes( x=ddd_level, y=population, fill=state  )) +
  geom_col() +
  ggtitle( 'Total Australian population living in LGAs with various DDD levels, 0-19 excluded')+
  facet_wrap(~age)

figure_103_age %>%
  mutate( ddd_level = cut( ddd, breaks=0:30*5 )) %>%
  ggplot( aes( x=ddd_level, y=population, fill=age  )) +
  geom_col() +
  ggtitle( 'Total Australian population living in LGAs with various DDD levels, 0-19 excluded')+
  facet_wrap(~state)

```



```{r figure_103_sex, message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient, by='pin') %>%
  filter( lga != '.' ) %>%
  select_and_standardise_ddd( standardise_over = qc( lga, sex )) %>% 
  { . } -> base_figure_103_sex

base_figure_103_sex %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               filter( lga != '.' ) %>%
               group_by( lga, state, sex ) %>%
               summarise( population = sum( population )), 
              by=c('lga', 'sex') ) %>% 
  filter( !endsWith(lga,  '99' )) %>% 
  { . } -> figure_103_sex


figure_103_sex %>%
  mutate( ddd_level = cut( ddd, breaks=0:20*7 )) %>%
  ggplot( aes( x=ddd_level, y=population, fill=state  )) +
  geom_col() +
  ggtitle( 'Total Australian population living in LGAs with various DDD levels')+
  facet_wrap(~sex)

figure_103_sex %>%
  mutate( ddd_level = cut( ddd, breaks=0:30*5 )) %>%
  ggplot( aes( x=ddd_level, y=population, fill=sex  )) +
  geom_col() +
  ggtitle( 'Total Australian population living in LGAs with various DDD levels')+
  facet_wrap(~state)

```

```{r replicate_AAOH_figure_103_actual,  message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient,  by='pin' ) %>%
   filter( lga != '.' & 
            !endsWith( lga, '99') &
            drug_type != 'benzodiazepine' &
            supply_date >= '2013-06-01' & supply_date<'2014-06-01') %>%
  group_by( lga ) %>%
  summarise( n = n()) %>%
  mutate( n = n *10) %>%
  inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               group_by( lga, state, lga_name ) %>%
               summarise( population = sum( population ))
             , by='lga' )  %>%
  mutate( rate = n/population * 100000 ) %>%
  arrange( rate ) %>%
  {.}->figure_103_actual

qplot( rate,data=figure_103_actual)
 
 ggplot(data = figure_103_actual) +
geom_dotplot(stackdir = 'center' ,   mapping = aes(x = rate))
  
```

```{r pbs precriptions for 2013 -14, message=FALSE, warning=FALSE}

df %>%
  inner_join( df_patient, by='pin') %>%
  filter( lga != '.' & 
            !endsWith( lga, '99') &
            drug_type != 'benzodiazepine' &
            supply_date >= '2013-06-01' & supply_date<'2014-06-01') %>%
  group_by( lga ) %>%
  summarise( n = n()) %>%
  mutate( n = n *10) %>%
  arrange( n ) %>% 
    inner_join( 
             df_population %>% 
               filter( supply_year==2013)  %>%
               group_by( lga, state, lga_name ) %>%
               summarise( population = sum( population ))
             , by='lga' ) 

```
  
```{r table_103, message=FALSE, warning=FALSE}
  
  base_figure_103 %>%
      inner_join(figure_103_actual , by='lga') %>%
      arrange(ddd )  %>%
      select(lga , lga_name , state , rate , n , ddd ) %>%
      rename (prescriptions = n) %>%
      {.}->figure_103_table

figure_103_table %>% 
  head(20) 

figure_103_table %>% 
  tail(20) 




  
  
```

  
  
  

